---
title: "Spatial Machine Learning"
author: "Justin Williams"
date: "2022-09-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages}
library(dotenv)
library(tidyverse)
library(sf)
library(tidycensus)
library(patchwork)
library(units)
library(leaflet)
library(mapview)
library(tigris)
options(tigris_use_cache = TRUE)
```

## Load data

Load in census data for NYC main racial demographics with **tidycensus** package.
```{r set-variables}
nyc_counties <- c("Bronx","Kings","New York","Queens","Richmond")
```

Racial data.

```{r load-racial-data}
get_acs(
  geography = "tract",
  variables = c(
    white = "B03002_003",
    black = "B03002_004",
    asian = "B03002_006",
    hispanic = "B03002_012"
  ), 
  state = "NY",
  county = nyc_counties,
  geometry = TRUE,
  year = 2019,
  key = Sys.getenv("CENSUS_API"))
```

### Median home variables for linear regression data

Search for variables

```{r load-variables}
v20 <- load_variables(2020, "acs5/profile", cache = TRUE)

View(v20)
```


ACS estimates acquired are:
  - `median_valueE`: The median home value of the Census tract (our outcome variable);

  - `median_roomsE`: The median number of rooms for homes in the Census tract;

  - `total_populationE`: The total population;

  - `median_ageE`: The median age of the population in the Census tract;

  - `median_year_builtE`: The median year built of housing structures in the tract;

  - `median_incomeE`: The median income of households in the Census tract;

  - `pct_collegeE`: The percentage of the population age 25 and up with a four-year college degree;

  - `pct_foreign_bornE`: The percentage of the population born outside the United States;

  - `pct_whiteE`: The percentage of the population that identifies as non-Hispanic white;
  
  - `pct_blackE`: The percentage of the population that identifies as non-Hispanic black;
  
  - `pct_hispanicE`: The percentage of the population that identifies as Hispanic, any race;
  
  - `pct_asianE`: The percentage of the population that identifies as Asian one race.

  - `percent_oohE`: The percentage of housing units in the tract that are owner-occupied.

```{r load-median-home-data}
# variable list
variables <- c(
  median_value = "B25077_001",
  median_rooms = "B25018_001",
  median_income = "DP03_0062",
  total_population = "B01003_001",
  median_age = "B01002_001",
  pct_college = "DP02_0068P",
  pct_foreign_born = "DP02_0094P",
  pct_white = "DP05_0077P",
  pct_black = "DP05_0078P",
  pct_hispanic = "DP05_0070P",
  pct_asian = "DP05_0044P",
  median_year_built = "B25037_001",
  percent_ooh = "DP04_0046P"
)

# get acs data
nyc_census_data <- get_acs(
  geography = "tract",
  variables = variables,
  state = "NY",
  county = nyc_counties,
  geometry = TRUE,
  output = "wide",
  year = 2020,
  key = Sys.getenv("CENSUS_API")) %>% 
  st_transform(2263)
```

View target variable

```{r target-variable}
(ggplot_med_income <- 
  ggplot(nyc_census_data, aes(fill = median_valueE)) +
  geom_sf(color = NA) + 
  scale_fill_viridis_c(labels = scales::label_dollar()) + 
  theme_void() + 
  labs(title = "NYC Median Home Value",
       subtitle = "(2016-2020 ACS Estimate)",
       fill = "Median home value ")) +
  theme(plot.title.position = 'plot',
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

# # save plot
# ggsave("./images/ggplot_med_inc.png", 
#        plot = ggplot_med_income,
#        bg = "white")
```

Try this with mapview

```{r mapview-version}
(mapview_nyc_med_home <- nyc_census_data %>% 
  mapview(
    zcol = "median_valueE",
    legend = T,
    layer.name = "NYC Median Home Value",
    alpha = 0.9,
    lwd=0.125,
    color = "white",
    na.alpha = 0.0))

# # save as png
# mapshot(mapview_nyc_med_home,
#         file = "./images/mapview_nyc_med_home.png")
```

Missing target values

```{r missing-target-value}
count(nyc_census_data[nyc_census_data$median_valueE == "NA",])$n
```

Histogram of `median_valueE` distribution.

```{r histogram}
(nyc_hist <- 
  nyc_census_data %>% 
  ggplot(aes(x = median_valueE)) + 
  geom_histogram(alpha = 0.5, fill = "navy", color = "navy",
                 bins = 100) + 
  theme_minimal() + 
  scale_x_continuous(labels = scales::label_dollar()) + 
  labs(x = "Median home value"))
```

Right skewed with a clustering of Census Tracts in the lower distribution and a long left tail in the expensive areas. This could lead to a violation of the normality principle let's transforming. 

### Transformation

  - try some type of transformation, sqrt?

```{r transform}
ggplot_med_home_sqrt <- nyc_census_data %>% 
  ggplot(aes(fill = sqrt(median_valueE))) + 
    geom_sf(color = NA) + 
    scale_fill_viridis_c() + 
    theme_void() + 
    labs(title = "NYC Median Home Income",
       subtitle = "(2016-2020 ACS Estimate)",
       fill = "Median home\nvalue (sqrt) ") +
    theme(plot.title.position = 'plot',
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

nyc_hist_sqrt <- nyc_census_data %>% 
  ggplot(aes(x = sqrt(median_valueE))) + 
  geom_histogram(alpha = 0.5, fill = "navy", color = "navy",
                 bins = 100) + 
  theme_minimal() + 
  scale_x_continuous() + 
  labs(x = "Median home value (sqrt)")

ggplot_med_home_sqrt + nyc_hist_sqrt
```

### Select only estimates, omit NA


```{r selectE-omit-NA}
(nyc_census_data_prepped <- nyc_census_data %>% 
  mutate(pop_density = as.numeric(set_units(total_populationE / st_area(.), "1/km2")),
         median_structure_age = 2020 - median_year_builtE) %>% 
  select(!ends_with("M")) %>% 
  rename_with(.fn = ~str_remove(.x, "E$")) %>%
  na.omit())
```

## Modeling

```{r first-model}
formula <- "sqrt(median_value) ~ median_rooms + median_income + pct_college + pct_foreign_born + pct_white + pct_black + pct_hispanic + pct_asian + median_age + percent_ooh + total_population + median_structure_age + pop_density"

model1 <- lm(formula = formula, data = nyc_census_data_prepped)

summary(model1)

```

